---
title: "EDA on second Data Pull (24-11-2019)"
---

## Prepare Notebook

```{r}
library(igraph)
library(magrittr)
library(lubridate)
library(networkD3)
library(tidytext)
library(tidyverse)
library(tm)
library(widyr)
library(wordcloud)
```

## Read Data

```{r}
raw_df <- readRDS(file = "../data/25_11_2019_twitter_21n.rds")
```

## EDA

Let us get a glimpse on the data. 

```{r}
glimpse(raw_df)
```

### Languages

**Warning:** We have tweets in japanese and arabic. We might be able to remove them using the `lang` feature. 

```{r}
raw_df %>% 
  filter(is.na(lang)) %>% 
  nrow()
```

```{r, fig.align="center"}
raw_df %>% 
  group_by(lang) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(lang = reorder(lang, n)) %>%  
  ggplot(mapping = aes(x = lang, y = n)) +
  geom_col() +
  coord_flip() +
  ggtitle(label = "Language Count")
```

Spanish is by far the most used language. 

### Timeframe 

Let us see the number of tweets over time. 

First, note that the `created_at` feature refers to UTC time. We need to substract 5 hours to get Colombian time. 
```{r}
raw_df$created_at[1] - (5*3600)
```


```{r}
raw_df %>% 
  mutate(created_at_round_hour = (created_at - (5*3600)) %>% round(units = "hour") %>% as.POSIXct()) %>% 
  group_by(created_at_round_hour) %>% 
  count() %>% 
  ggplot(mapping = aes(x = created_at_round_hour, y = n)) +
  geom_line() +
  ggtitle(label = "Number of Tweets (All) per Hour")
```

There is a remarkable high peack 

```{r}
raw_df %>% 
  mutate(created_at_round_hour = (created_at - (5*3600)) %>% round(units = "hour") %>% as.POSIXct()) %>% 
  group_by(created_at_round_hour) %>% 
  count() %>% 
  arrange(-n) %>% 
  head(10)
```

The most active hours where on the 21st of November between 20:00 and 23:00. There were many events around this time period which could have contributed to this: cacerolazo and the media spread on chaos in Bogota and Cali. We need to understand this better. 

```{r}
raw_df %>% 
  mutate(created_at_round_hour = (created_at - (5*3600)) %>% round(units = "hour") %>% as.POSIXct()) %>% 
  group_by(created_at_round_hour, is_retweet) %>% 
  count() %>% 
  ggplot(mapping = aes(x = created_at_round_hour, y = n, color = is_retweet)) +
  geom_line() +
  ggtitle(label = "Number of Tweets (All) per Hour")
```

### Location

```{r, fig.align="center"}
raw_df %>% 
  mutate(location = case_when(location == "" ~ "no-tag", TRUE ~ str_to_lower(location))) %>% 
  group_by(location) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(n = n /sum(n)) %>% 
  arrange(- n) %>% 
  head(20) %>% 
  mutate(lang = reorder(location, n)) %>%  
  ggplot(mapping = aes(x = lang, y = n)) +
  geom_col() +
  coord_flip() +
  ggtitle(label = "Location Count")
```

### Hashtags

```{r}
hashtags_vect <- raw_df %>% 
  filter(! is.na(hashtags)) %>% 
  pull(hashtags) %>% 
  unlist()

hashtags_df <- tibble(hashtag = hashtags_vect) %>% 
  mutate(hashtag = str_to_lower(string = hashtag)) %>% 
  count(hashtag) %>% 
  arrange(-n)
```

```{r}
hashtags_df %>% mutate(n = n /sum(n)) %>% 
  top_n(n = 30, wt = n) %>% 
  mutate(hashtag = reorder(hashtag, n)) %>%
  ggplot(aes(x = hashtag, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  ggtitle(label = 'Hashtag Count')
```

```{r}
wordcloud(
  words = hashtags_df$hashtag, 
  freq = hashtags_df$n, 
  min.freq = 500, 
  colors = brewer.pal(8, "Dark2"))

```

### Top - Tweets

Top (direct) retweeted people:

```{r}
raw_df %>% 
  filter(is_retweet == FALSE) %>% 
  group_by(screen_name) %>% 
  summarise(retweet_count = sum(retweet_count)) %>% 
  ungroup() %>% 
  arrange(- retweet_count) %>% 
  head(30) %>% 
  mutate(screen_name = reorder(screen_name, retweet_count)) %>%
  ggplot(aes(x = screen_name, y = retweet_count)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  ggtitle(label = 'Top Retweeted Users')
```

Let us get the text of the most retweeted mentions:

```{r}
raw_df %>% 
  filter(is_retweet == FALSE) %>% 
  select(screen_name, text, retweet_count) %>% 
  arrange(- retweet_count) %>% 
  head(10)
```


### Retweet Network 

Let us define a retweet weighted directed network as follows:

- Nodes: 
  - Source = `screen_name` 
  - Target = `retweet_screen_name`
  
- Edges: Two nodes are conected if there is at least `threshold` number of retweets. 

- Weight: number of retweets. 

For visualization purposes: Let `threshold` > 0. We can filter edges which have weight less than `threshold`. 

First, we construct a data frame. 

```{r}
# Construct retweet network. 
retweets_df <- raw_df %>% 
  filter(is_retweet == TRUE) %>% 
  select(screen_name, retweet_screen_name) %>% 
  count(screen_name, retweet_screen_name)
```

We now construct the network from the dataframe above. 

```{r}
retweets_graph <- graph_from_data_frame(d = retweets_df, directed = TRUE)
```


```{r}
# We compute the in/out weighted degree for the retweet graph. 
compute_weighted_deg_df <- function (retweets_graph) {
  
  retweets_node_df <- tibble(
    screen_name = V(retweets_graph)$name
  )
  
  tweets_out_df <- retweets_df %>% 
    select(screen_name, n) %>% 
    group_by(screen_name) %>% 
    summarise(tweets_out = sum(n))
  
  tweets_in_df <- retweets_df %>%
    select(retweet_screen_name, n) %>% 
    group_by(retweet_screen_name) %>% 
    summarise(tweets_in = sum(n))
  
  retweets_node_df %<>% 
    left_join(y = tweets_out_df, by = c("screen_name")) %>% 
    left_join(y = tweets_in_df, by = c("screen_name" = "retweet_screen_name")) %>% 
    replace_na(list(tweets_in = 0, tweets_out = 0))
  
  return(retweets_node_df)
}

retweets_node_df <- compute_weighted_deg_df(retweets_graph)
```

```{r}
retweets_node_df %>% arrange(- tweets_in)
```

We can sort with respect to `tweets_in`, which measures users whose tweets where retweeted the most. 

```{r}
retweets_node_df %>% arrange(- tweets_in)
```

Next, we sort with respect to `tweets_out`, which measures users who retweeted the most. 

```{r}
retweets_node_df %>% arrange(- tweets_out)
```

```{r}
V(retweets_graph)$tweets_in <- retweets_node_df$tweets_in
V(retweets_graph)$tweets_out <- retweets_node_df$tweets_out
```


- Visualization

```{r}
# Set weight threshold. 
threshold <- 13

red_retweets_df <- raw_df %>% 
  filter(is_retweet == TRUE) %>% 
  select(screen_name, retweet_screen_name) %>% 
  count(screen_name, retweet_screen_name) %>% 
  filter(n >= threshold)

red_retweets_graph <- graph_from_data_frame(d = red_retweets_df, directed = TRUE)

red_retweets_node_df <- compute_weighted_deg_df(red_retweets_graph)

V(red_retweets_graph)$tweets_in <- red_retweets_node_df$tweets_in
V(red_retweets_graph)$tweets_out <- red_retweets_node_df$tweets_out
```

```{r}
# Compute the weight shares.
E(red_retweets_graph)$width <- E(red_retweets_graph)$n/max(E(red_retweets_graph)$n)

# Create networkD3 object.
network_d3 <- igraph_to_networkD3(g = red_retweets_graph)
# Define node size.
network_d3$nodes %<>% mutate(Degree = 0.01*V(red_retweets_graph)$tweets_in)
# Degine color group (I will explore this feature later).
network_d3$nodes %<>% mutate(Group = 1)
# Define edges width. 
network_d3$links$Width <- 10*E(red_retweets_graph)$width

network_d3_viz <- forceNetwork(
  Links = network_d3$links, 
  Nodes = network_d3$nodes, 
  arrows = TRUE,
  Source = 'source', 
  Target = 'target',
  NodeID = 'name',
  Group = 'Group', 
  opacity = 0.9,
  Value = 'Width',
  Nodesize = 'Degree', 
  # We input a JavaScript function.
  linkWidth = JS("function(d) { return Math.sqrt(d.value); }"), 
  fontSize = 12,
  zoom = TRUE, 
  opacityNoHover = 1, 
)

saveNetwork(
  network = network_d3_viz, 
  file = paste0(getwd(), "/eda_second_data_pull_output/retweet_network.html"), 
  selfcontained = TRUE
)
```

### Text Analysis

- Text Cleaning

```{r}
tweets_corpus <- VCorpus(
  VectorSource(
    raw_df %>% 
      filter(lang == "es") %>% 
      mutate(text = str_to_lower(text)) %>% 
      pull(text)
  )
)
```

```{r}
rm_pattern <- content_transformer( 
  function (x, pattern) str_remove_all(string = x, pattern = pattern)
)

url_pattern <- " ?(f|ht)(tp)(s?):(//)?\\S+"

clean_tweets_corpus <- tweets_corpus %>% 
  tm_map(rm_pattern, "/") %>% 
  tm_map(rm_pattern, "\\|") %>% 
  tm_map(rm_pattern, url_pattern) %>% 
  tm_map(stripWhitespace) %>% 
  tm_map(PlainTextDocument)

clean_tweets_df <- tibble(
  clean_text = sapply(X = clean_tweets_corpus, FUN = function(x) x$content)
)
```

```{r}
clean_tweets_df %>% head(100)
```

- Word Count

```{r}
stopwords_df <- tibble(word = stopwords::stopwords(language = "es"))

words_df <- clean_tweets_df %>% 
  unnest_tokens(output = word, input = clean_text) %>% 
  anti_join(y = stopwords_df, by = "word")

word_count <- words_df %>% 
  count(word) %>% 
  arrange(- n)
```

```{r}
word_count %>% 
  head(30) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  ggtitle(label = "Word Count")
```


```{r, fig.width=8, fig.align="center"}
wordcloud(
  words = word_count$word, 
  freq = word_count$n, 
  min.freq = 2500, 
  colors = brewer.pal(8, "Dark2"))
```

- 2-Gram Network

```{r}
bigram_count <- clean_tweets_df %>% 
  unnest_tokens(input = clean_text, output = bigram, token = "ngrams", n = 2) %>% 
  separate(col = bigram, into = c("word1", "word2"), sep = " ") %>% 
  filter(! word1 %in% stopwords_df$word) %>% 
  filter(! word2 %in% stopwords_df$word) %>% 
  filter(! is.na(word1)) %>% 
  filter(! is.na(word2)) %>% 
  count(word1, word2, sort = TRUE)
  
```

```{r}
bigram_threshold <- 3500

bigram_network <- bigram_count %>%
  filter(n > bigram_threshold) %>%
  graph_from_data_frame(directed = FALSE)
```

```{r, fig.width=10, fig.height=10,  fig.align="center"}
pdf(paste0(getwd(),"/eda_second_data_pull_output/bigram_network.pdf")) 

V(bigram_network)$degree <- degree(graph = bigram_network)

E(bigram_network)$width <- E(bigram_network)$n/max(E(bigram_network)$n)

plot(
  bigram_network, 
  vertex.size = 0.3*V(bigram_network)$degree,
  vertex.color = "lightblue",
  vertex.label.color = "black", 
  vertex.label.cex = 0.3, 
  vertex.label.dist = 0.2,
  edge.color = "gray", 
  edge.width = 1*E(bigram_network)$width,
  main = "Bigram Count Network",
  alpha = 50, 
  layout = layout_with_fr(bigram_network), 
)

dev.off() 
```



