---
title: "EDA on first Data Pull (19-11-2019)"
---

## Prepare Notebook

```{r}
library(igraph)
library(magrittr)
library(lubridate)
library(networkD3)
library(tidyverse)
```

## Read Data

```{r}
raw_df <- readRDS(file = "../data/19_11_2019_twitter_21n.rds")
```

## EDA

Let us get a glimpse on the data. 

```{r}
glimpse(raw_df)
```

### Languages

**Warning:** We have tweets in japanese and arabic. We might be able to remove them using the `lang` feature. 

```{r}
raw_df %>% 
  filter(is.na(lang)) %>% 
  nrow()
```

```{r, fig.align="center"}
raw_df %>% 
  group_by(lang) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(lang = reorder(lang, n)) %>%  
  ggplot(mapping = aes(x = lang, y = n)) +
  geom_col() +
  coord_flip() +
  ggtitle(label = "Language Count")
```

Spanish is by far the most used language. 

### Timeframe 

```{r}
raw_df %>% 
  mutate(created_at_round_hour = created_at %>% round(units = "hour") %>% as.POSIXct()) %>% 
  group_by(created_at_round_hour) %>% 
  count() %>% 
  ggplot(mapping = aes(x = created_at_round_hour, y = n)) +
  geom_line() +
  ggtitle(label = "Number of Tweets (All) per Hour")
```

```{r}
raw_df %>% 
  mutate(created_at_round_hour = created_at %>% round(units = "hour") %>% as.POSIXct()) %>% 
  group_by(created_at_round_hour, is_retweet) %>% 
  count() %>% 
  ggplot(mapping = aes(x = created_at_round_hour, y = n, color = is_retweet)) +
  geom_line() +
  ggtitle(label = "Number of Tweets (All) per Hour")
```

We do not see a big increase in organic tweets. 

### Hashtags

```{r}
hashtags_vect <- raw_df %>% 
  filter(! is.na(hashtags)) %>% 
  pull(hashtags) %>% 
  unlist()

hashtags_df <- tibble(hashtag = hashtags_vect) %>% 
  mutate(hashtag = str_to_lower(string = hashtag)) %>% 
  count(hashtag) %>% 
  arrange(-n)
```

```{r, fig.height=5}
hashtags_df %>% mutate(n = n /sum(n)) %>% 
  top_n(n = 40, wt = n) %>% 
  mutate(hashtag = reorder(hashtag, n)) %>%
  ggplot(aes(x = hashtag, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  ggtitle(label = 'Hashtag Count')
```

### Retweet Network 

Let us define a retweet weighted directed network as follows:

- Nodes: 
  - Source = `screen_name` 
  - Target = `retweet_screen_name`
  
- Edges: Two nodes are conected if there is at least `threshold` number of retweets. 

- Weight: number of retweets. 

For visualization purposes: Let `threshold` > 0. We can filter edges which have weight less than `threshold`. 

First, we construct a data frame. 

```{r}
# Construct retweet network. 
retweets_df <- raw_df %>% 
  filter(is_retweet == TRUE) %>% 
  select(screen_name, retweet_screen_name) %>% 
  count(screen_name, retweet_screen_name)
```

We now construct the network from the dataframe above. 

```{r}
retweets_graph <- graph_from_data_frame(d = retweets_df, directed = TRUE)
```


```{r}
# We compute the in/out weighted degree for the retweet graph. 
compute_weighted_deg_df <- function (retweets_graph) {
  
  retweets_node_df <- tibble(
    screen_name = V(retweets_graph)$name
  )
  
  tweets_out_df <- retweets_df %>% 
    select(screen_name, n) %>% 
    group_by(screen_name) %>% 
    summarise(tweets_out = sum(n))
  
  tweets_in_df <- retweets_df %>%
    select(retweet_screen_name, n) %>% 
    group_by(retweet_screen_name) %>% 
    summarise(tweets_in = sum(n))
  
  retweets_node_df %<>% 
    left_join(y = tweets_out_df, by = c("screen_name")) %>% 
    left_join(y = tweets_in_df, by = c("screen_name" = "retweet_screen_name")) %>% 
    replace_na(list(tweets_in = 0, tweets_out = 0))
  
  return(retweets_node_df)
}

retweets_node_df <- compute_weighted_deg_df(retweets_graph)
```

We can sort with respect to `tweets_in`, which measures users whose tweets where retweeted the most. 

```{r}
retweets_node_df %>% arrange(- tweets_in)
```

Next, we sort with respect to `tweets_out`, which measures users who retweeted the most. 

```{r}
retweets_node_df %>% arrange(- tweets_out)
```

```{r}
V(retweets_graph)$tweets_in <- retweets_node_df$tweets_in
V(retweets_graph)$tweets_out <- retweets_node_df$tweets_out
```


- Visualization

```{r}
# Set weight threshold. 
threshold <- 5

red_retweets_df <- raw_df %>% 
  filter(is_retweet == TRUE) %>% 
  select(screen_name, retweet_screen_name) %>% 
  count(screen_name, retweet_screen_name) %>% 
  filter(n >= threshold)

red_retweets_graph <- graph_from_data_frame(d = red_retweets_df, directed = TRUE)

red_retweets_node_df <- compute_weighted_deg_df(red_retweets_graph)

V(red_retweets_graph)$tweets_in <- red_retweets_node_df$tweets_in
V(red_retweets_graph)$tweets_out <- red_retweets_node_df$tweets_out
```

```{r}
# Compute the weight shares.
E(red_retweets_graph)$width <- E(red_retweets_graph)$n/max(E(red_retweets_graph)$n)

# Create networkD3 object.
network_d3 <- igraph_to_networkD3(g = red_retweets_graph)
# Define node size.
network_d3$nodes %<>% mutate(Degree = 0.01*V(red_retweets_graph)$tweets_in)
# Degine color group (I will explore this feature later).
network_d3$nodes %<>% mutate(Group = 1)
# Define edges width. 
network_d3$links$Width <- 10*E(red_retweets_graph)$width

network_d3_viz <- forceNetwork(
  Links = network_d3$links, 
  Nodes = network_d3$nodes, 
  arrows = TRUE,
  Source = 'source', 
  Target = 'target',
  NodeID = 'name',
  Group = 'Group', 
  opacity = 0.9,
  Value = 'Width',
  Nodesize = 'Degree', 
  # We input a JavaScript function.
  linkWidth = JS("function(d) { return Math.sqrt(d.value); }"), 
  fontSize = 12,
  zoom = TRUE, 
  opacityNoHover = 1, 
)

saveNetwork(network = network_d3_viz, file = "retweet_network.html", selfcontained = TRUE)
```

