---
title: "Retweet Network"
---

## Prepare Notebook

```{r, warning=FALSE,  message=FALSE}
knitr::opts_knit$set(warning=FALSE, message=FALSE)

set.seed(seed = 42)

if(!require(igraph)) install.packages("igraph",repos = "http://cran.us.r-project.org")
library("igraph")
if(!require(magrittr)) install.packages("magrittr",repos = "http://cran.us.r-project.org")
library("magrittr")
if(!require(lubridate)) install.packages("lubridate",repos = "http://cran.us.r-project.org")
library("lubridate")
if(!require(tidytext)) install.packages("tidytext",repos = "http://cran.us.r-project.org")
library("tidytext")
if(!require(tidyverse)) install.packages("tidyverse",repos = "http://cran.us.r-project.org")
library("tidyverse")

source(file = "../R/retweet_network.R")
```


## Read Data

```{r}
data_df <- readRDS(file = "../data/data_df.rds")
```

## Construct Complete Retweet Network 

```{r}
retweet_network_df <- get_retweet_network_df(data_df = data_df)
retweet_network <- get_retweet_network(retweet_network_df = retweet_network_df)
```

## In/Out Degree

```{r}
retweets_deg_df <- get_retweets_deg_df(retweets_network = retweet_network )
```

We can sort with respect to `tweets_in`, which measures users whose tweets where retweeted the most. We select the top ten retweeted people to make a more readable graph later on. 

```{r}
retweets_deg_df %>% 
  arrange(- tweets_in) %>% 
  head(10)
```

Next, we sort with respect to `tweets_out`, which measures users who retweeted the most. 

```{r}
retweets_deg_df %>% 
  arrange(- tweets_out) %>% 
  head(10)
```

---
**NOTE**

I am not sure this is very useful information
---

## Visualize Retweet Network 

```{r}
red_retweet_network_df <- get_retweet_network_df(data_df = data_df, weight_threshold = 11)
red_retweet_network <- get_retweet_network(retweet_network_df = red_retweet_network_df)
retweets_deg_df <- get_retweets_deg_df(retweets_network = red_retweet_network )

top_retweeted <- retweets_deg_df %>% 
  arrange(- tweets_in) %>% 
  head(20) %>% 
  pull(screen_name)

top_retweeted_names <- ifelse(
  test = V(red_retweet_network)$name %in% top_retweeted, 
  yes = V(red_retweet_network)$name, 
  no = ""
)
```

---
**NOTE**

Q: Why did we choose 11 as the threshold?
A: Visualization purposes. 
---

```{r}
pdf("../images/retweet_network.pdf") 

plot(
  red_retweet_network , 
  vertex.color = "lightblue",
  vertex.frame.color = "lightblue",
  # Scale node size by degree.
  vertex.size = (5E-4)*(V(red_retweet_network)$tweets_in + 1),
  vertex.label.color = "black", 
  vertex.label.cex = (4.3E-2)*((V(red_retweet_network)$tweets_in + 1)^(1/5)), 
  vertex.label.dist = 0.0,
  vertex.label = top_retweeted_names,
  edge.color = "gray", 
  # Set edge width proportional to the weight relative value.
  edge.width = (1E-3)*E(red_retweet_network)$weight_scaled,
  edge.arrow.size = 0.03,
  main = '21N - Twitter', 
  sub = "Retweet Network", 
  alpha = 50
)

dev.off()
```
---
**NOTE**

Q: Do we need the proportional width for the edges? I can't really see any difference in the figure. 
A: This is WIP, might be useful for other plots. 
---

---
**NOTE**

Q: I wonder if we can make a static version of the network? I think the benefit of having a dynamic network is outweighed by the computational cost of having it running in the background of the browser. 
A: The current network is static (pdf)

Q: Can we remove the names of the users that retweeted the most retweeted individuals? I think it's interesting to see that, for example, Felix de Bedut got retweeted a lot, but I am not really interested in who did it, unless it was someone else that got retweeted a lot. What do you think? 
A: Good idea! Lets do it on a different PR. 